\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[shortlabels]{enumitem}
%\setcounter{secnumdepth}{0}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{fancyhdr, lastpage}
\pagestyle{fancy}
\fancyhf{}
\lhead{Daniel Standage}
\chead{BCB 567, 11:00am T/Th}
\rhead{Lecture Notes: August 24, 2010}
%\cfoot{Page \thepage{} of \protect\pageref*{LastPage}}
\usepackage{varioref}
\labelformat{equation}{(#1)}
\usepackage[colorlinks,linkcolor=blue]{hyperref}

\newenvironment{mitemize}
{
  \begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}}{\end{itemize}
}

\newenvironment{menumerate}
{
  \begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}}{\end{enumerate}
}


\begin{document}

\section*{Algorithms}
Why do we need algorithms? We will look at an example.

\subsection*{Rock Game}
The following is a description of the rock game.
\begin{mitemize}
\item There are two players, Alice and Bob
\item There are two piles of rocks, each with ten rocks
\item Rules are as follows
  \begin{mitemize}
  \item In a turn you may either take a rock from one of the piles or a rock from each of the piles
  \item The player who takes the last stone(s) wins
  \end{mitemize}
\end{mitemize}

Given this setup, is there a winning strategy for Alice or Bob? This \textbf{winning strategy} may also be the key to understanding more complex, less trivial problems. For example, the winning strategy to this game can easily be applied to the alignment of two sequences.

\subsection*{What is an algorithm?}
\begin{mitemize}
\item A well-defined computational procedure that takes some value(s) as INPUT and produces some value(s) as OUTPUT
\item In mathematical terms, an algorithm is a function that maps every element in some input space to an element in some output space
\end{mitemize}

One example of an algorithm is the sorting problem. The input to this algorithm is a sequence of integers $A = (a_1, a_2, ..., a_n)$. The output is some permutation $A'$ of $A$ such that $a_1' \leq a_2' \leq ... \leq a_n'$.

\end{document}